\chapter{Data Presentation and Analysis}
In this section, the collected traffic data is presented and analyzed to understand traffic patterns and evaluate the performance of the AI-based traffic management system.

\section{Data Presentation}

The data collected from sensors and cameras include vehicle counts, traffic speed, congestion levels, and signal timings at different intersections and time intervals. This data is organized into tables and graphs for better visualization.

\begin{itemize}
    \item \textbf{Vehicle Count:} Number of vehicles passing through an intersection per minute.
    \item \textbf{Average Speed:} Average speed of vehicles at various points.
    \item \textbf{Congestion Levels:} Degree of traffic buildup measured by waiting time and queue length.
    \item \textbf{Signal Timings:} Duration of green, yellow, and red signals adjusted dynamically.
\end{itemize}

\section{Data Analysis}

The analysis focuses on how the AI system improves traffic flow by adjusting signals in real-time. Key observations include:

\begin{itemize}
    \item Reduction in average waiting time at traffic signals during peak hours.
    \item Improvement in average vehicle speed through intersections.
    \item Decrease in congestion levels compared to traditional fixed-time signal systems.
    \item Prioritization of emergency vehicles, resulting in faster clearance.
\end{itemize}

Statistical tools and visualization techniques such as line graphs, bar charts, and heat maps are used to illustrate the improvements and validate the effectiveness of the AI model.


