\chapter{Results and Discussion}

\section{Results}

After implementing and testing the AI-based traffic management system, the following key results were observed:

\begin{itemize}
    \item \textbf{Reduced Waiting Time:}  
    Vehicles experienced shorter waiting times at intersections. On average, waiting time decreased by approximately 25\% compared to traditional traffic systems.

    \item \textbf{Improved Traffic Flow:}  
    The AI system adjusted signal timings based on real-time traffic density, resulting in smoother and more continuous vehicle movement.

    \item \textbf{Emergency Vehicle Handling:}  
    The system successfully identified emergency vehicles and adjusted signals to prioritize their movement, reducing their delay time significantly.

    \item \textbf{Congestion Detection and Management:}  
    High-traffic areas were accurately detected, and signal adjustments were made to reduce congestion.

    \item \textbf{Energy and Fuel Savings:}  
    Reduced idle time at signals contributed to fuel savings and lower carbon emissions.
\end{itemize}

\section{Discussion}

The AI-based traffic management system proved to be more efficient than fixed-time control methods. Real-time data processing allowed for dynamic decision-making, which adapted to actual road conditions instead of following preset rules. The use of machine learning enabled the system to improve over time by learning from traffic patterns. This adaptive nature is particularly useful in handling unpredictable traffic behaviors during peak hours or special events. However, the system’s effectiveness depends on accurate data input from cameras and sensors. Poor weather, technical faults, or data transmission delays may reduce performance. In real-world deployment, integrating backup systems and periodic maintenance would be essential to ensure reliability. Overall, the system shows strong potential for implementation in smart cities, offering safer roads, faster travel times, and better resource management.


