\chapter*{Appendices}

\addcontentsline{toc}{chapter}{Appendices}

\section*{Appendix A: Sample Vehicle Detection Logs}

The following table shows a sample extract of the raw data collected by the vehicle detection module. It records vehicle counts per direction, signal states, and timestamps at 10-second intervals.

\begin{table}[H]
\centering
\caption*{Vehicle Detection Log Sample}
\begin{tabular}{|c|c|c|c|c|c|}
\hline
\textbf{Timestamp} & \textbf{North} & \textbf{South} & \textbf{East} & \textbf{West} & \textbf{Green Signal} \\
\hline
10:00:00 & 15 & 8  & 12 & 5  & North \\
10:00:10 & 10 & 9  & 14 & 7  & East \\
10:00:20 & 6  & 15 & 8  & 10 & South \\
\hline
\end{tabular}
\end{table}

\vspace{0.5cm}

\section*{Appendix B: Core Python Script for Vehicle Detection}

Below is a snippet from the Python script that performs real-time vehicle detection using the YOLOv8 model and ONNX Runtime:

\begin{verbatim}
import onnxruntime as ort
import cv2
import numpy as np

# Load YOLOv8 ONNX model
session = ort.InferenceSession("yolov8.onnx")

def preprocess_frame(frame):
    # Resize and normalize frame
    resized = cv2.resize(frame, (640, 640))
    img = resized.transpose(2, 0, 1) / 255.0
    img = img[np.newaxis, ...].astype(np.float32)
    return img

def detect_vehicles(frame):
    input_tensor = preprocess_frame(frame)
    outputs = session.run(None, {'images': input_tensor})
    # Process outputs to extract bounding boxes and class IDs
    # ...
    return detections
\end{verbatim}

\vspace{0.5cm}

\section*{Appendix C: Backend API Endpoints}

The backend server exposes REST and WebSocket endpoints to handle data ingestion and communication with the frontend and AI modules.

\begin{itemize}
    \item \texttt{POST /api/vehicle-count} - Receives JSON payloads containing vehicle detection counts.
    \item \texttt{GET /api/signal-status} - Provides the current traffic signal state.
    \item \texttt{WS /ws/live} - WebSocket endpoint streaming real-time traffic updates to the frontend dashboard.
\end{itemize}

\vspace{0.5cm}

\section*{Appendix D: Sample JSON Payload for Vehicle Counts}

\begin{verbatim}
{
  "timestamp": "2025-07-06T10:00:00Z",
  "counts": {
    "north": 15,
    "south": 8,
    "east": 12,
    "west": 5
  },
  "vehicle_types": {
    "car": 25,
    "motorcycle": 10,
    "bus": 3,
    "truck": 2
  },
  "green_signal": "north"
}
\end{verbatim}

\vspace{0.5cm}

\section*{Appendix E: System Deployment Configuration}

Details of the Google Cloud Platform (GCP) VM setup used for deployment:

\begin{itemize}
    \item OS: Ubuntu 22.04 LTS
    \item CPU: 4 vCPUs
    \item RAM: 8 GB
    \item GPU: NVIDIA Tesla T4
    \item Docker installed for containerizing backend and AI modules
    \item Ktor backend daemonized with systemd
    \item Python AI scripts managed with Supervisor
\end{itemize}

\vspace{0.5cm}

\section*{Appendix F: Additional Experiment Logs}

This section contains extended logs of system performance metrics such as latency, detection accuracy over multiple test runs, and system uptime monitoring, available upon request or in the supplementary materials.

---

\textit{The appendices provide additional technical details to complement the main body of the report and support reproducibility and further development of the AI-Based Traffic Management System.}